|

### 2021-01-22 supplementary_information.Rmd
Update supplementary_information.Rmd

@@

# Theoretical predictions for longitudinal gradients of local species richness

Local biodiversity patterns have been extensively studied in rivers, illuminating typical patterns observed in nature. The most common pattern is a downstream increase of local species richness[@mcdowallDiadromyAssemblyRestoration1996; @mcdowallFightingFlowDownstreamupstream1998; @rahelFishAssemblagesHabitat1991; @teruiThreeEcologicalFactors2016]. However, recent empirical and theoretical studies also showed 'reversed patterns', in which local species richness decreases downstream[@bsemerHeadwatersAreCritical2013; @harveyDisturbanceReversesClassic2018]. We predicted the longitudinal gradient of local species richness to confirm that our simulation scenarios are capable of reproducing previously observed patterns of local species richness. We considered 32 simulation scenarios comprising four landscape and eight ecological scenarios as described in the maintext (a set of parameters is described in **Table SXX**). Under each simulation scenario, we generated 10 replicates of branching networks with fixed parameters of ecosystem size ($N_p = 100$) and complexity ($P_b = 0.5$). We evaluated the relationship between local species richness and the number of upstream habitat patches, a proxy for the longitudinal position of a habitat patch.

@@

>>>>

# simulation data
dat_sim <- read_csv(here::here('/theory/result/result_sim2020-11-10.csv'))

@@

# Sensitivity analysis

# Longitudinal gradients of local species richness

Longitudinal gradients of local species richness have been extensively studied in rivers, illuminating typical patterns observed in nature. The most common pattern is a downstream increase of local species richness[@mcdowallDiadromyAssemblyRestoration1996; @mcdowallFightingFlowDownstreamupstream1998; @rahelFishAssemblagesHabitat1991; @teruiThreeEcologicalFactors2016]. However, recent empirical and theoretical studies also showed 'reversed patterns', in which local species richness decreases downstream[@bsemerHeadwatersAreCritical2013; @harveyDisturbanceReversesClassic2018]. We predicted the longitudinal gradient of local species richness to confirm that our simulation scenarios are capable of reproducing previously observed patterns of local species richness. We considered 32 simulation scenarios comprising four landscape and eight ecological scenarios as described in the maintext (a set of parameters is described in **Table SXX**). Under each simulation scenario, we generated 10 replicates of branching networks with fixed parameters of ecosystem size ($N_p = 100$) and complexity ($P_b = 0.5$). We evaluated the relationship between local species richness and the number of upstream habitat patches, a proxy for the longitudinal position of a habitat patch.

@@

## Table S1 Simulation parameter

Values and interpretation of parameters used in our simulation model. Refer to the maintext for model details.

```{r}

dat_sim %>% 
  select(sd_env_source,
         sd_env_lon,
         sd_env,
         rho,
         phi,
         niche_cost,
         max_alpha,
         theta,
         p_dispersal
         ) %>% 
  summarise(across(everything(), unique)) %>% 
  mutate(r0 = 4) %>% 
  pivot_longer(cols = everything(),
               names_to = "Parameter",
               values_to = "Value") %>% 
  group_by(Parameter) %>% 
  summarize(Value = list(unique(Value))) %>% 
  mutate(id = case_when(Parameter == 'sd_env_source' ~ 'a',
                        Parameter == 'sd_env_lon' ~ 'b',
                        Parameter == 'sd_env' ~ 'c',
                        Parameter == 'rho' ~ 'd',
                        Parameter == 'phi' ~ 'e',
                        Parameter == 'niche_cost' ~ 'f',
                        Parameter == 'max_alpha' ~ 'g',
                        Parameter == 'theta' ~ 'h',
                        Parameter == 'p_dispersal' ~ 'i',
                        Parameter == 'r0' ~ 'j',
                        TRUE ~ as.character(Parameter)),
         Interpretation = case_when(Parameter == 'sd_env_source' ~ 'Environmental variation at headwaters',
                                    Parameter == 'sd_env_lon' ~ 'Degree of local environmental noise',
                                    Parameter == 'sd_env' ~ 'Temporal environmental variability',
                                    Parameter == 'rho' ~ 'Strength of spatial autocorrelation in mean environmental condition',
                                    Parameter == 'phi' ~ 'Extent of spatial autocorrelation in temporal environmental variation',
                                    Parameter == 'niche_cost' ~ 'Cost of a wider niche',
                                    Parameter == 'max_alpha' ~ 'Maximum value of interspecific competition coefficient',
                                    Parameter == 'theta' ~ 'Rate parameter of an exponential dispersal kernel',
                                    Parameter == 'p_dispersal' ~ 'Dispersal probability',
                                    Parameter == 'r0' ~ 'Maxiumum reproductive rate',
                                    TRUE ~ as.character(Parameter)),
         Parameter = case_when(Parameter == 'sd_env_source' ~ '$\\sigma_{h}$',
                               Parameter == 'sd_env_lon' ~ '$\\sigma_{l}$',
                               Parameter == 'sd_env' ~ '$\\sigma_{z}$',
                               Parameter == 'rho' ~ '$\\rho$',
                               Parameter == 'phi' ~ '$\\phi$',
                               Parameter == 'niche_cost' ~ '$\\nu$',
                               Parameter == 'max_alpha' ~ '$\\alpha_{max}$',
                               Parameter == 'theta' ~ '$\\theta$',
                               Parameter == 'p_dispersal' ~ '$p_{d}$',
                               Parameter == 'r0' ~ '$r_{0,i}$',
                               TRUE ~ as.character(Parameter))
         ) %>% 
  arrange(id) %>% 
  select(-id) %>% 
  kable(format = 'markdown')
